A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-04 paper.Rmd
section 5.4 revised

## Effect of shape of non-linearity

A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern. According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials. But the chance of detecting the "triple-U" shape drops significantly compared to other shapes. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969]. However, the "triple-U" shape constructed from the Hermite polynomials use power up to 18. If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be only slight lower than the power curve of "M" shape. The recommendation of the polynomial power for the RESET test should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.

For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "triple-U" shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "S" shape. This implies, unlike the conventional test, the visibility of the shape do not strictly follow the degree of the polynomials.

```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.'}

>>>>

## Effect of amount of non-linearity

The order of the polynomial is a primary factor contributing to the pattern produced by the non-linearity model. Figure \ref{fig:poly-power-uniform-j} explores the relationship between polynomial order and power of the tests. The conventional tests have higher power for lower orders of Hermite polynomials, and it drops substantially for the "triple-U" shape. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969]. However, the "triple-U" shape is constructed from the Hermite polynomials using power up to 18. If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be closer to other power curves. This illustrates the sensitivity of the conventional test to the parameter choice, and highlights a limitation that it helps to know the data generating process to set the parameters for the test, which is unrealistic. However, we examined this in more detail (see Appendix) and found that there is no harm for setting the parameter higher than four on the tests' operation for lower order polynomial shapes. Using a parameter value of six, instead of four, yields higher power regardless of generating process, and would be recommended.

For visual tests, we expect the "U" shape to be detected more readily, followed by the "S", "M" and "triple-U" shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves mostly align with these expectations, except for the "M" shape, which is as easy to be detected as the "S" shape. This suggests a benefit of  the visual test, the knowing the shape ahead of time is not needed for its application.

```{r poly-power-uniform-j, fig.height = 4, fig.cap = 'The effect of the order of the polynomial on the power of conventional and visual tests. Deeper colour indicates higher order. The default RESET tests under-performs significantly in detecting the "triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test, but this higher than the recommended value.'}

### 2023-04-04 paper.tex
section 5.4 revised

\hypertarget{effect-of-shape-of-non-linearity}{%
\subsection{Effect of shape of
non-linearity}\label{effect-of-shape-of-non-linearity}}

A primary factor contributes to the non-linearity model is the shape of
the non-linearity pattern. According to Figure
\ref{fig:poly-power-uniform-j}, conventional tests have higher power in
testing shapes constructed from lower orders of Hermite polynomials. But
the chance of detecting the ``triple-U'' shape drops significantly
compared to other shapes. To understand why this is, one needs to return
to the way the RESET test is applied. It requires a parameter indicating
degree of fitted values to test for, and the recommendation is to
generically use four \citep{ramsey_tests_1969}. However, the
``triple-U'' shape constructed from the Hermite polynomials use power up
to 18. If the RESET test had been applied using a higher power no less
than six, the power curve of ``triple-U'' shape will be only slight
lower than the power curve of ``M'' shape. The recommendation of the
polynomial power for the RESET test should be revised, perhaps. This
illustrates the sensitivity of conventional testing to the parameters,
and it also points to a limitation that one needs to know the data
generating process in order to set the parameters for the test.

For visual tests, based on the orders of the Hermite polynomials, we
expect the ``U'' shape will be the easiest one to be detected by
subjects followed by the ``S'', ``M'' and ``triple-U'' shape. From
Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power
curves are mostly aligned with our expectation, except for the ``M''
shape, which is as easy to be detected as the ``S'' shape. This implies,
unlike the conventional test, the visibility of the shape do not
strictly follow the degree of the polynomials.

@@

\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours. The default RESET tests under-perform significantly in detecting "triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test. But this means the order needs to be known prior to testing.}\label{fig:poly-power-uniform-j}
\end{figure}

>>>>

\hypertarget{effect-of-amount-of-non-linearity}{%
\subsection{Effect of amount of
non-linearity}\label{effect-of-amount-of-non-linearity}}

The order of the polynomial is a primary factor contributing to the
pattern produced by the non-linearity model. Figure
\ref{fig:poly-power-uniform-j} explores the relationship between
polynomial order and power of the tests. The conventional tests have
higher power for lower orders of Hermite polynomials, and it drops
substantially for the ``triple-U'' shape. To understand why this is, one
needs to return to the way the RESET test is applied. It requires a
parameter indicating degree of fitted values to test for, and the
recommendation is to generically use four \citep{ramsey_tests_1969}.
However, the ``triple-U'' shape is constructed from the Hermite
polynomials using power up to 18. If the RESET test had been applied
using a higher power no less than six, the power curve of ``triple-U''
shape will be closer to other power curves. This illustrates the
sensitivity of the conventional test to the parameter choice, and
highlights a limitation that it helps to know the data generating
process to set the parameters for the test, which is unrealistic.
However, we examined this in more detail (see Appendix) and found that
there is no harm for setting the parameter higher than four on the
tests' operation for lower order polynomial shapes. Using a parameter
value of six, instead of four, yields higher power regardless of
generating process, and would be recommended.

For visual tests, we expect the ``U'' shape to be detected more readily,
followed by the ``S'', ``M'' and ``triple-U'' shape. From Figure
\ref{fig:poly-power-uniform-j}, it can be observed that the power curves
mostly align with these expectations, except for the ``M'' shape, which
is as easy to be detected as the ``S'' shape. This suggests a benefit of
the visual test, the knowing the shape ahead of time is not needed for
its application.

@@

\caption{The effect of the order of the polynomial on the power of conventional and visual tests. Deeper colour indicates higher order. The default RESET tests under-performs significantly in detecting the "triple-U" shape. To achieve a similar power as other shapes, a higher order polynomial parameter needs to be used for the RESET test, but this higher than the recommended value.}\label{fig:poly-power-uniform-j}
\end{figure}
